Forecasting the power of Higher Order Weak Lensing Statistics with automatically differentiable simulations

### 2022-02-14 paper/paper.tex

%\begin{equation}\label{displacement}
 %   \textbf{S}=4\pi G \bar{\rho}\alpha_{PM}/H_0^2 \nabla(
  % \hat{\textbf{O}}_l(k)
   %  \hat{\textbf{O}}_s(k)
    %\nabla^{-2}\delta),
%\end{equation}
%where $\hat{\textbf{O}}_l(k)= \exp{(\frac{k^2}{k^4_l})}$ and $\hat{\textbf{O}}_s(k)= \exp{(-\frac{k^4}{k^4_s}})$ are the low and high pass filter introduced to remove the long range force and to reduce the numerical effect induced by the mesh resolution. \\
PGD introduces control parameters that can be tuned comparing the matter power spectrum to a reference power spectrum.

@@

In Figure \ref{fig:comp_dens_field} , we show the resulting density field when this operation is applied. We show a comparison of the density field image for a given redshift when the PGD correction is applied and when not \denise{write better this}. We can see that the structures are not modified but simply sharpened.
\denise{image size and position}
% \begin{figure}
%     \centering
%     \includegraphics[width=\columnwidth]{paper/figures/densitymap.png}
%     \caption{Comparison of density planes at same redshift with and without the PGD correction.}
%     \label{fig:comp_dens_field}
% \end{figure}

Figure \ref{fig:comp_kmap} shows the same result on the convergence map.
\denise{image size and position}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{paper/figures/kmap_compgd.png}
    \caption{Comparison of the final convergence map with and without the PGD correction.}
    \label{fig:comp_kmap}
\end{figure}

In Figure \ref{fig:ps_comp} we compare the low resolution DLL convergence power spectrum complemented by the PGD scheme to a conventional DLL simulation with the same resolution. Both the outputs are compared to the angular power spectrum obtained from the $\kappa$TNG only dark matter simulation. We can conclude that PGD scheme allows to achieve higher accuracy at the same computational cost.
\denise{image size and position}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{paper/figures/ps_comp.png}
    \caption{Comparison of the angular power spectrum from the final convergence map with and without the PGD correction against $\kappa-$TNG only Dark-matter simulations.}
    \label{fig:ps_comp}
\end{figure}

\subsection{Summary statistics}
In this work, in order to extract the cosmological information from the simulated $\kappa-$maps, we used tree statistics: the angular power spectrum, the peak counts and the $l_{1\text{norm}}$.

@@

\subsubsection{Lensing Angular Power Spectrum}
Second order statistics, both in the form of shear 2-point correlation function $\xi_{\pm}(\theta)$, or its counterpart in Fourier space, the angular power spectrum $C_{\ell}$, have been widely used to extract the cosmological information from weak lensing mass maps. \
In the Limber approximation, the angular power spectrum of the convergence field can be computed as:
\begin{equation}
    C_{\kappa}(\ell)=\frac{9\Omega_m^2H_0^4}{4c^4}
    \int_0^{\chi_{lim}} \text{d}\chi 
    \frac{g^2(\chi)}{a^2(\chi)}P_{\delta}
    	\left( k=\frac{\ell}{f_K(\chi)},\chi \right),
\end{equation}
where $P_{\delta}$ defines the matter power spectrum of the density contrast.
In this paper, the angular power spectrum is computed from the noisy smoothed convergence maps: to each mock $\kappa$map we add Gaussian noise and than convolved it with a Gaussian smoothing kernel with a smoothing size $\theta=3.6$. We perform the angular power spectrum analysis in a range of $\ell=[300,3000]$ according to the requirement specified in  \cite{mandelbaum2018lsst}. 
For each map, we compute the power spectrum by using our own code implemented in the Tensorflow framework and we take the average over 2000 maps realisation.

>>>>

%\begin{equation}\label{displacement}
 %   \textbf{S}=4\pi G \bar{\rho}\alpha_{PM}/H_0^2 \nabla(
  % \hat{\textbf{O}}_l(k)
   %  \hat{\textbf{O}}_s(k)
    %\nabla^{-2}\delta),
%\end{equation}
%where $\hat{\textbf{O}}_l(k)= \exp{(\frac{k^2}{k^4_l})}$ and $\hat{\textbf{O}}_s(k)= \exp{(-\frac{k^4}{k^4_s}})$ are the low and high pass filter introduced to remove the long range force and to reduce the numerical effect induced by the mesh resolution. \\
PGD introduces control parameters that can be tuned by comparing the matter power spectrum to a reference power spectrum.

@@

Figure \ref{fig:pkhalofit_comp} shows the matter power spectrum at $z=1.69$ of our simulations before and after applying the PGD correction compared to analytic Halofit predictions (\denise{cita}(Smith et al. 2003; Takahashi et al. 2012)).
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{paper/figures/pkhalofit_comp.png}
    \caption{
    Upper panel: Comparison of the matter power spectrum of FlowPM before and after the PGD correction with reference matter power spectrum \textit{Halofit} at redshift $z = 1.69$. Lower panel: Fractional differences between FlowPM and Halofit.
    }
    \label{fig:pkhalofit_comp}
\end{figure}

In Figure \ref{fig:comp_kmap} we show an example of our convergence map at $ z= 1.034$, as well as the same map when the PGD correction is not applied. 
We can see that the structures are not modified but simply sharpened.
\begin{figure*}
    \centering
    \includegraphics[width=0.9\textwidth]{paper/figures/kmap_compgd.png}
    \caption{ \textbf{Left panel}: An example of our convergence map at source redshift $z= 1.034$.
    \textbf{Right panel}: Same convergence map when the PGD correction is not applied.}
    \label{fig:comp_kmap}
\end{figure*}

In Figure \ref{fig:ps_comp} in the upper panel, we present the DLL angular power spectrum complemented by the PGD scheme to a conventional DLL simulation with the same resolution. Both the outputs are compared to the analytic Halofit prediction for the angular power spectrum. In the lower panel the fractional differences between the our maps and the analytic Halofit prediction.  Both the power spectra and ratios are averaged over $N = 100$ realisations.
We can conclude that PGD scheme allows to achieve higher accuracy at the same computational cost.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{paper/figures/clshalofit_comp.png}
    \caption{
    \textbf{Upper panel}: Angular power spectra for 3 different source redshifts from
our DLL maps. \textbf{Lower panel}: Fractional differences between the
DLL and Halofit power spectra for 3 source redshifts. The power spectra and ratios are means over 100 map realisations. The two vertical lines define the range of $\ell$ between $[300,3000]$.}
    \label{fig:ps_comp}
\end{figure}

\subsection{Summary statistics}
In this work, in order to extract the cosmological information from the simulated $\kappa-$maps, we use tree summary statistic for weak lensing observable: the angular power spectrum, the `starlet peak counts' and the `starlet $l_{1\text{norm}}$'.

@@

\subsubsection{Lensing Angular Power Spectrum}
Second order statistics, both in the form of shear 2-point correlation function $\xi_{\pm}(\theta)$, or its counterpart in Fourier space, the angular power spectrum $C_{\ell}$, have been widely used to extract the cosmological information from weak lensing mass maps. \
In the Limber approximation, the angular power spectrum of the convergence field can be computed as:
\begin{equation}
    C_{\kappa}(\ell)=\frac{9\Omega_m^2H_0^4}{4c^4}
    \int_0^{\chi_{lim}} \text{d}\chi 
    \frac{g^2(\chi)}{a^2(\chi)}P_{\delta}
    	\left( k=\frac{\ell}{f_K(\chi)},\chi \right),
\end{equation}
where $P_{\delta}$ defines the matter power spectrum of the density contrast.

In this paper, the angular power spectrum is computed from the noisy smoothed convergence maps: to each mock $\kappa$map we add Gaussian noise and than convolved it with a Gaussian smoothing kernel.

To understand how accurate our simulations are, and to validate the choice of the smoothing size to apply, we tried to reproduce the summary statistics from the only Dark $\kappa-TNG$ simulations. 
 We show the results from the angular power spectrum comparison for 3 source redshifts in Figure \ref{fig:ps_comp}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{paper/figures/clshalofit_comp.png}
    \caption{
    \textbf{Upper panel}: Angular power spectra for 3 different source redshifts from
our DLL maps. \textbf{Lower panel}: Fractional differences between the
DLL and Halofit power spectra for 3 source redshifts. The power spectra and ratios are means over 100 map realisations. The two vertical lines define the range of $\ell$ between $[300,3000]$.}
    \label{fig:ps_comp}
\end{figure}
% with a smoothing size $\theta=3.6$. We perform the angular power spectrum analysis in a range of $\ell=[300,3000]$ according to the requirement specified in  \cite{mandelbaum2018lsst}.

For each map, we compute the power spectrum by using our own code implemented in the Tensorflow framework and we take the average over 2000 maps realisation.
